Bayesian analysis of the NESTA study of interventions against  verbal aggression online

### 2021-09-10 bayesianAnalysis/bayesianReport.Rmd
comparing models

\begin{align*}
\mathsf{AdiffS} & \sim \textsf{Norm}(\mu, \sigma)\\
mu_i & = \alpha + \beta_{\mathsf{ADS}}[\mathsf{group}_i]\times \mathsf{ADS} + \beta_{\mathsf{group}_i}  +
 \beta_{\mathsf{IC}}[\mathsf{group}_i]\times \mathsf{IC} + \\
 & + \beta_{\mathsf{ADSIC}}\times \mathsf{ADS} \times \mathsf{IC} + \beta_{\mathsf{CBS}}[\mathsf{group}_i] \times \mathsf{CBS}\\
 \alpha & \sim \textsf{Norm}(0,.3)\\
\beta_{\mathsf{ADS}}[\mathsf{group}_i] & \sim \textsf{Norm}(0,.3)\\
\beta_{\mathsf{group}_i} & \sim \textsf{Norm}(0,.3)\\
\beta_{\mathsf{IC}}[\mathsf{group}_i] & \sim \textsf{Norm}(0,.3)\\
 \beta_{\mathsf{ADSIC}} & \sim \textsf{Norm}(0,.3)\\
 \beta_{\mathsf{CBS}}[\mathsf{group}_i]& \sim \textsf{Norm}(0,.3)\\
\end{align*}

@@

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", results = "hide"}
mu <- link(InteractionsModelDiff)
mu_mean <- apply( mu , 2 , mean )
mu_resid <- summaries$AdiffS - mu_mean
ggplot()+geom_density(aes(x = mu_resid))+theme_tufte()+
  ggtitle("Residuals are approximately normally distributed")
```
\normalsize

@@

>>>>

\begin{align*}
\mathsf{AdiffS} & \sim \textsf{Norm}(\mu, \sigma)\\
\mu_i & = \alpha + \beta_{\mathsf{ADS}}[\mathsf{group}_i]\times \mathsf{ADS} + \beta_{\mathsf{group}_i}  +
 \beta_{\mathsf{IC}}[\mathsf{group}_i]\times \mathsf{IC} + \\
 & + \beta_{\mathsf{ADSIC}}\times \mathsf{ADS} \times \mathsf{IC} + \beta_{\mathsf{CBS}}[\mathsf{group}_i] \times \mathsf{CBS}\\
 \alpha & \sim \textsf{Norm}(0,.3)\\
\beta_{\mathsf{ADS}}[\mathsf{group}_i] & \sim \textsf{Norm}(0,.3)\\
\beta_{\mathsf{group}_i} & \sim \textsf{Norm}(0,.3)\\
\beta_{\mathsf{IC}}[\mathsf{group}_i] & \sim \textsf{Norm}(0,.3)\\
 \beta_{\mathsf{ADSIC}} & \sim \textsf{Norm}(0,.3)\\
 \beta_{\mathsf{CBS}}[\mathsf{group}_i]& \sim \textsf{Norm}(0,.3)\\
\end{align*}

@@

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", results = "hide"}
mu <- link(InteractionsModelDiff)
mu_mean <- apply( mu , 2 , mean )
mu_resid <- summaries$AdiffS - mu_mean
ggplot()+geom_density(aes(x = mu_resid))+theme_tufte()+
  ggtitle("Residuals are approximately normally distributed")+xlab("residuals")
```
\normalsize

@@

# Model selection

How did we get to this fairly complicated model though? Once preliminary causal considerations guided our restrictions on variable selection, we proceed by building models of increasing complexity, and comparing them in terms of Widely Acceptable Information Criterion. The models differ mostly in the underlying linear formulae. For computational ease we will here use quadratic approximations, while in the final analysis we will deploy Hamiltionian Monte Carlo. The names are meant to decode the model structure: the predictors are listed before dashes, whereas interactions are listed after dashes.

\begin{align}
\tag{Null}  \mu_i & = \alpha\\
\tag{ADS}  \mu_i & = \alpha + \beta_{\mathsf{ADS}}\times \mathsf{ADS}\\
\tag{ADSIC}  \mu_i & = \alpha + \beta_{\mathsf{ADS}}\times \mathsf{ADS} +    \beta_{\mathsf{IC}}\times \mathsf{IC}\\
\tag{IT}  \mu_i & = \beta_{\mathsf{group}[i]} \\
\tag{ADSIT} \mu_i & = \alpha + \beta_{\mathsf{ADS}}\times \mathsf{ADS} +  \beta_{\mathsf{group}[i]}\\
\tag{ADSITIC} \mu_i & = \alpha + \beta_{\mathsf{ADS}}\times \mathsf{ADS} +  \beta_{\mathsf{group}[i]} +    \beta_{\mathsf{IC}}\times \mathsf{IC}\\
\tag{ADSITIC-ADSIC} \mu_i & = \alpha + \beta_{\mathsf{ADS}}\times \mathsf{ADS} +  \beta_{\mathsf{group}[i]} +    \beta_{\mathsf{IC}}\times \mathsf{IC} + \beta_{\mathsf{ADSIC}}\times \mathsf{ADS} \times \mathsf{IC}\\
\tag{ADSITIT-ADSIC-ADSIT} \mu_i & = \alpha + \beta_{\mathsf{ADS}}[\mathsf{group}_i]\times \mathsf{ADS} +  \beta_{\mathsf{group}[i]} +  \\ \nonumber & +  \beta_{\mathsf{IC}}\times \mathsf{IC} + \beta_{\mathsf{ADSIC}}\times \mathsf{ADS} \times \mathsf{IC}\\
\tag{ADSIT-ADSIT}   \mu_i &  = \alpha + \beta_{\mathsf{ADS}}[\mathsf{group}_i] \times
 \mathsf{ADS} + \beta_{\mathsf{group}[i]} \\
\tag{ADSITIC-ADSIT-ICIT-ADSIC}   \mu_i &  = \alpha  + \beta_{\mathsf{group}[i]} +  
\beta_{\mathsf{IC}}[\mathsf{group}_i] \times \mathsf{IC}  + \beta_{\mathsf{ADSIC}}\times \mathsf{ADS} \times \mathsf{IC}\\
\tag{ADSITICCBS-ITIC-ADSIC} \mu_i &  = \alpha  + \beta_{\mathsf{group}[i]} +  
\beta_{\mathsf{ADS}}[\mathsf{group}_i] \times
 \mathsf{ADS}  + \beta_{\mathsf{IC}}[\mathsf{group}_i] \times \mathsf{IC}  + \\ & \nonumber  \beta_{\mathsf{ADSIC}}\times \mathsf{ADS} \times \mathsf{IC} +
 \beta_{\mathsf{group}[i]} +  
\beta_{\mathsf{CBS}} \times \mathsf{CBS} \\
\tag{Final} \mu_i & = \alpha + \beta_{\mathsf{ADS}}[\mathsf{group}_i]\times \mathsf{ADS} + \beta_{\mathsf{group}_i}  +  \beta_{\mathsf{IC}}[\mathsf{group}_i]\times \mathsf{IC} + \\
 & + \beta_{\mathsf{ADSIC}}\times \mathsf{ADS} \times \mathsf{IC} + \beta_{\mathsf{CBS}}[\mathsf{group}_i] \times \mathsf{CBS} \nonumber \end{align}

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
null <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu ~ dnorm (0,0.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries  
)

ADS <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <-  a + bADS * ADS,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,0.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSIC <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <-  a + bADS * ADS+ bIC * IC,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,0.3),
    bIC ~ dnorm(0,0.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

IT <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <-  bIT[groupID] ,
    bIT[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSIT <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS * ADS +  bIT[groupID],
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITIC <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS * ADS +  bIT[groupID] + bIC * IC,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITIC_ADSIC <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS * ADS +  bIT[groupID] + bIC * IC + bADSIC * ADS * IC,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITIT_ADSIC_ADSIT <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC * IC + bADSIC * ADS * IC,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSIT_ADSIT <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] ,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    #bADSIC ~ dnorm(0,.5),
    bIT[groupID] ~ dnorm(0,.3),
    #bIC ~ dnorm(0,.5),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITIC_ADSIT_ICIT_ADSIC <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC +
      bADSIC * ADS * IC,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITICCBS_ITIC_ADSIC <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
      bADSIC * ADS * IC+ bCBS *CBS,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCBS ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

Final <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
      bADSIC * ADS * IC+ bCBS[groupID] *CBS,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCBS[groupID] ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
comparison<- compare(null,ADS,ADSIC,IT,ADSIT,ADSITIC,ADSITIC_ADSIC,ADSITIT_ADSIC_ADSIT,ADSIT_ADSIT,ADSITIC_ADSIT_ICIT_ADSIC,ADSITICCBS_ITIC_ADSIC,Final)
plot(comparison)
```
\normalsize
